### 2023-04-05 main.tex
i

\paragraph{Proof of \eqref{edgeDualRealtion}.}To show \eqref{edgeDualRealtion} we need two 'ingredients'. First the existence of a similarity transformation between the Hamiltonian $H_{x,y}$ defined in \eqref{edgeHamiltonian} and its transposed. As we will show, this similarity transformation is $R_{x}$ defined in \eqref{Rmatrix}. Second, the possibility of finding a symmetry for the Hamiltonian \eqref{edgeHamiltonian}. Exploiting the fact that $H_{x,y}$ is a linear function of the coproduct of the second Casimir of $gl(N)$, (see \eqref{hamiltonianCasimir}) this symmetry is $\exp{(E^{x})}\exp{(E^{y})}$ with $E^{x}$ defined in \eqref{EquationEx}. \\
Equation \eqref{transpositionPropertyR} implies that 
\begin{equation}\label{transpositionPropertyH}
    H_{x,y}^{T}=\left(R_{x}R_{y}\right)H_{x,y}\left(R_{x}R_{y}\right)^{-1}
As a consequence of \eqref{hamiltonianCasimir} we look for a symmetry of  $\Delta(C_{2})$.

%We now search for a symmetry of $H_{x,y}$. % Given $A,B\in gl(N)$, we say that $A$ is a symmetry of $B$ if 
%\begin{equation}
%    [A,B]=0
%\end{equation}
Since the coproduct is a Lie algebra homomorphism we have that $\forall A,B\in gl(N)$
\begin{equation}\label{symmetryCoproduct}
    [A,B]=0\quad \Longrightarrow\quad \left[\Delta(A),\Delta(B)\right]=0
\end{equation}
%Since $H_{x,y}$ is a linear function of the coproduct of the second Casimir, as a consequence of \eqref{symmetryCoproduct}, it is enough to identify a symmetry of $C_{2}$. But
 $C_{2}$ is in the center of $gl(N)$, and therefore it commutes with all the element of the the algebra. Considering 
\begin{equation}\label{equationE}
    E=\sum_{a=2}^{N}E_{a1}
\end{equation}
to obtain a product structure of the elements \eqref{dualityElements} of the duality matrix, we chose as a symmetry 
\begin{equation}
    S=\exp{(\Delta(E))}
\end{equation}
%Since $[C_{2},E]=0$, we have that 
%\begin{equation}\label{commutationSC}
    %\left[S,\Delta(C_{2})\right]=0
%\end{equation}
For fixed $(x,y)\in \mathcal{E}$ we introduce
\begin{equation}
    S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. As a consequence of \eqref{symmetryCoproduct}, we have that 
\begin{equation}\label{symmetryH}
    \left[S_{x,y},H_{x,y}\right]=0
\end{equation}
i.e. $S_{x,y}$ is a symmetry of $H_{x,y}$. \\ Finally, we obtain that 
\begin{equation}
    \begin{split}
        H_{x,y}^{T}D&=(R_{x}R_{y})H_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{\widehat{y}}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{\widehat{y}}\right)H_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=
        DH_{x,y}
    \end{split}
\end{equation}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 %\textbf{Reamrk}: it is important to notice that the elements of the duality matrix \eqref{dualityElements} are well defined only if at each site $x\in V$ and for every species $k\in \{2,\ldots,N\}$ the number of dual particles $r_{x}^{a}$ is lower or equal than the number of original particles $n_{a}^{x}$. This implies that the dynamics of the dual process is simpler, because described by a lower number of particles. By the way, it is possible to perform computations similar to the one made in this proof choosing instead of  \eqref{equationE} {\color{red} $E=\sum_{a=2}^{N}E_{1a}$ or $E=\sum_{a=2}^{N}E_{aa}$.}
 %{\color{blue} That's confusing. What is E?}

% However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we need to transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $\widehat{x}$. % that we connect to every site $x\in V$ of the graph $G$. \\
 Considering $A,B\in gl(N)$, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-B)}A\exp{(B)}=A-\left[B,A\right]+\frac{1}{2!}\left[B,\left[B,A\right]\right]-\frac{1}{3!}\left[B,\left[B,\left[B,A\right]\right]\right]+\ldots
 \end{equation}
In the followin we need the application of the Hadamard formula with $B=E$ defined in \eqref{equationE} and with $A$ equal to some of the generator $E_{ab}$ of the Lie algebra. Therefore we compute 
\begin{enumerate}
    \item for $A=E_{b 1}$ with $b\in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT_El1}
        \exp{(-E)}E_{b 1}\exp{(E)}=E_{b 1}
    \end{equation}
    because $E_{b 1}$ commutes with $E$
    \item for $A=E_{bb}$ with $b \in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT_Ell}
        \exp{(-E)}E_{b b}\exp{(E)}=E_{b b}+E_{b 1}
    \end{equation}
Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
       \left[E, E_{bb}\right]=\sum_{a=2}^{N}\left(E_{ab}\delta_{b 1}-E_{b 1}\delta_{ab}\right)=-E_{b1}
    \end{equation}
    Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT_Ell}.
    \item for $A=E_{1b}$ with $b\in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT-E1l}
        \exp{(-E)}E_{1 b}\exp{(E)}=E_{1b}+E_{11}-\sum_{a=2}^{N}\left(E_{a1}+E_{ab}\right)
    \end{equation}
  %\com{Why indices have names $b, k, j, a, b$? Fix one set for the generators}

    Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
[E,E_{1b}]=\sum_{a=2}^{N}\left(E_{ab}\delta_{11}-E_{11}\delta_{ab}\right)= (\sum_{a=2}^{N}E_{ab} ) -E_{11};
\end{equation}
and 
\begin{equation}
\begin{split}
\left[E,[E,E_{1b}]\right]&=\sum_{b=2}^{N}\sum_{a=2}^{N}\left[E_{b1},E_{ab}\right]-\sum_{b=2}^{N}\left[E_{b1},E_{11}\right]
\\&=
\sum_{a,b=2}^{N}\left(E_{bb}\delta_{a1}-E_{a1}\delta_{bb}\right)-\sum_{b=2}^{N}\left(E_{b1}\delta_{11}-E_{11}\delta_{b1}\right)
\\=&
-2\sum_{a=2}^{N}E_{a1};
\end{split}
\end{equation}
Inserting the above commutators in \eqref{HadamardFormula}, we obtain \eqref{HT-E1l}.
\item for $A=E_{ab}$ with $a,b \in \{2,\ldots,N\}$ we obtain 
\begin{equation}\label{HT-Ekl}
    \exp{(-E)}E_{a b}\exp{(E)}=E_{ab}+E_{a 1}
\end{equation}
   Indeed, using \eqref{eq:comgl} we have 
\begin{equation}
[E,E_{ab}]=\sum_{c=2}^{N}\left(E_{cb}\delta_{a1}-E_{a1}\delta_{c b}\right)=-E_{a1};
\end{equation}
Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT-Ekl}.
\item for $A=E_{11}$ we obtain 
\begin{equation}\label{HT-E11}
    \exp{(-E)}E_{11}\exp{(E)}=E_{11}-\sum_{a=2}^{N}E_{a1}
\end{equation}
  Indeed, using \eqref{eq:comgl} we have

>>>>

\paragraph{Proof of \eqref{edgeDualRealtion}.}To show this relation we need two 'ingredients'. First the existence of a similarity transformation between the Hamiltonian $H_{x,y}$  and its transposed. As we will show, this similarity transformation is $R_{x}R_{y}$.  Second, the possibility of finding a symmetry for the edge Hamiltonian. Exploiting \eqref{hamiltonianCasimir}, we will search for a symmetry of the second Casimir $C_{2}$. As we will see, this symmetry is $\sum_{a=2}^{N}E_{a1}$ \\
Using \eqref{transpositionPropertyR} we obtain

\begin{equation}\label{transpositionPropertyH}
    H_{x,y}^{T}=\left(R_{x}R_{y}\right)H_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the symilarity tranfsormation between $H_{x,y}$ and its transposed. \\
As already pointed out, $H_{x,y}$ is a linear function of $\Delta(C_{2})$. This implies that symmetries of  $\Delta(C_{2})$ are also symmetries of  $H_{x,y}$. Moreover, the coproduct is a Lie algebra homomorphism, i.e. $\forall A,B\in gl(N)$
\begin{equation}\label{symmetryCoproduct}
    [A,B]=0\quad \Longrightarrow\quad \left[\Delta(A),\Delta(B)\right]=0
\end{equation}
This tells us that it is enough to find a symmetry of $C_{2}$. This element belongs to the center of $gl(N)$, and therefore it commutes with all the element of this algebra. Consider
\begin{equation}\label{equationE}
    E=\sum_{a=2}^{N}E_{a1}
\end{equation}
To obtain a product structure of the elements of the duality matrix, we introduce
\begin{equation}
    S=\exp{(\Delta(E))}
\end{equation}
Finally, for fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
    S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. This operator $S^{x,y}$ satisfies
\begin{equation}\label{symmetryH}
    \left[S_{x,y},H_{x,y}\right]=0
\end{equation}
i.e. $S_{x,y}$ is a symmetry of $H_{x,y}$. \\ Exploiting these considerations we write
\begin{equation}
    \begin{split}
        H_{x,y}^{T}D&=(R_{x}R_{y})H_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{\widehat{y}}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{\widehat{y}}\right)H_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=
        DH_{x,y}
    \end{split}
\end{equation}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 %\textbf{Reamrk}: it is important to notice that the elements of the duality matrix \eqref{dualityElements} are well defined only if at each site $x\in V$ and for every species $k\in \{2,\ldots,N\}$ the number of dual particles $r_{x}^{a}$ is lower or equal than the number of original particles $n_{a}^{x}$. This implies that the dynamics of the dual process is simpler, because described by a lower number of particles. By the way, it is possible to perform computations similar to the one made in this proof choosing instead of  \eqref{equationE} {\color{red} $E=\sum_{a=2}^{N}E_{1a}$ or $E=\sum_{a=2}^{N}E_{aa}$.}
 %{\color{blue} That's confusing. What is E?}

% However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we need to transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $\widehat{x}$. % that we connect to every site $x\in V$ of the graph $G$. \\
 Considering $A,B\in gl(N)$, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-B)}A\exp{(B)}=A-\left[B,A\right]+\frac{1}{2!}\left[B,\left[B,A\right]\right]-\frac{1}{3!}\left[B,\left[B,\left[B,A\right]\right]\right]+\ldots
 \end{equation}
In the following we need the application of the Hadamard formula with $B=E$ defined in \eqref{equationE} and with $A$ equal to some of the generator $E_{ab}$ of the Lie algebra. Therefore we compute 
\begin{enumerate}
    \item for $A=E_{b 1}$ with $b\in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT_El1}
        \exp{(-E)}E_{b 1}\exp{(E)}=E_{b 1}
    \end{equation}
    because $E_{b 1}$ commutes with $E$
    \item for $A=E_{bb}$ with $b \in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT_Ell}
        \exp{(-E)}E_{b b}\exp{(E)}=E_{b b}+E_{b 1}
    \end{equation}
Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
       \left[E, E_{bb}\right]=\sum_{a=2}^{N}\left(E_{ab}\delta_{b 1}-E_{b 1}\delta_{ab}\right)=-E_{b1}
    \end{equation}
    Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT_Ell}.
    \item for $A=E_{1b}$ with $b\in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT-E1l}
        \exp{(-E)}E_{1 b}\exp{(E)}=E_{1b}+E_{11}-\sum_{a=2}^{N}\left(E_{a1}+E_{ab}\right)
    \end{equation}
  %\com{Why indices have names $b, k, j, a, b$? Fix one set for the generators}

    Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
[E,E_{1b}]=\sum_{a=2}^{N}\left(E_{ab}\delta_{11}-E_{11}\delta_{ab}\right)= (\sum_{a=2}^{N}E_{ab} ) -E_{11};
\end{equation}
and 
\begin{equation}
\begin{split}
\left[E,[E,E_{1b}]\right]&=\sum_{c=2}^{N}\sum_{a=2}^{N}\left[E_{c1},E_{ab}\right]-\sum_{c=2}^{N}\left[E_{c1},E_{11}\right]
\\&=
\sum_{a,c=2}^{N}\left(E_{cb}\delta_{a1}-E_{a1}\delta_{cb}\right)-\sum_{c=2}^{N}\left(E_{c1}\delta_{11}-E_{11}\delta_{c1}\right)
\\=&
-2\sum_{a=2}^{N}E_{a1};
\end{split}
\end{equation}
Inserting the above commutators in \eqref{HadamardFormula}, we obtain \eqref{HT-E1l}.
\item for $A=E_{ab}$ with $a,b \in \{2,\ldots,N\}$ we obtain 
\begin{equation}\label{HT-Ekl}
    \exp{(-E)}E_{a b}\exp{(E)}=E_{ab}+E_{a 1}
\end{equation}
   Indeed, using \eqref{eq:comgl} we have 
\begin{equation}
[E,E_{ab}]=\sum_{c=2}^{N}\left(E_{cb}\delta_{a1}-E_{a1}\delta_{c b}\right)=-E_{a1};
\end{equation}
Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT-Ekl}.
\item for $A=E_{11}$ we obtain 
\begin{equation}\label{HT-E11}
    \exp{(-E)}E_{11}\exp{(E)}=E_{11}-\sum_{a=2}^{N}E_{a1}
\end{equation}
  Indeed, using \eqref{eq:comgl} we have
